### 2011-11-21 doc/theses/vitasek/appendix-c-trace.tex
Thesis: merging Kesy, fixing singletons and duplex.

The first section deals with system information. Please note that some of these characteristics cannot be easily obtained programmatically and are thus stored in the source code as constants.\\
To obtain GLPK information, the program parses the first line of standard output produced by running \code{glpsol -v}. It tries to guess whether it's the Cygwin version by looking at the path to the binary.

@@

To be able to reconstruct the progress of the metaheuristic, the next section contains CSV \nomenclature{CSV}{Comma Separated Values} formatted data for each iteration. Each row contains the time in milliseconds, quality of the incumbent solution and the number of its AMs.

@@

And finally, it is important to know what is the ID/IDREF set recommended by this experiment run. Thus the log is concluded by a CSV formatted list of element - attribute name pairs to be included in the ID and IDREF set, respectively.

>>>>

The first section deals with system information. Note that some of~these char\-ac\-ter\-is\-tics cannot be easily obtained programmatically and are thus stored in the source code as constants.\\
To obtain GLPK information, the program parses the first line of standard output produced by running \code{glpsol -v}. It tries to guess whether it's the Cygwin version by looking at the path to the binary.

@@

To be able to reconstruct the progress of the metaheuristic, the next section contains CSV \nomenclature{CSV}{Comma Separated Values} formatted data for each iteration. Each row contains the time in~milliseconds, quality of the incumbent solution and the number of its AMs.

@@

And finally, it is important to know what is the ID/IDREF set recommended by this experiment run. Thus the log is concluded by a CSV formatted list of~element - attribute name pairs to be included in the ID and IDREF set, respectively.

### 2011-11-21 doc/theses/vitasek/chapter-3-mip.tex
Thesis: merging Kesy, fixing singletons and duplex.

In this chapter we introduce a new approach to finding maximum ID sets. First, we transform the problem formulation to maximum weighted IS problem formulation. Then we transform this into a MIP formulation, and demonstrate how this can be solved using a solver such as GLPK \cite{glpk}. \nomenclature{GLPK}{GNU Linear Programming Kit}
We will continue by applying heuristical approaches to improve the performance of the process.

@@

An advantage of using \textit{Branch \& Bound} is that while traversing the search tree it finds intermittent, sub-optimal solutions. It is thus possible to limit the total search time and instead of the optimum take the best solution found so far.\\

We will now demonstrate the full process of finding the optimal ID set of an example XML file using GLPK.

@@

The first three edges are due to the type collision ($y$), the last one is due to $\iota(M_{y}^{a}) \cap \iota(M_{z}^{a}) = \{1\}$. The graph $G$ constructed in this way is in Figure \ref{image-mip-is-graph}.

\begin{figure}
  \caption{IS Representation Graph}
  \label{image-mip-is-graph}
  \centering
	\includegraphics[width=.25\textwidth]{images/is-representation}
\end{figure}

The next step is the MIP formulation. We do not need to translate from the IS formulation, as the translation from ID set formulation is straightforward, too. For each AM $m$ there will be one binary variable $x_{name(m)}$. Objective function coefficients in vector $\mathbf{c}$ will be weights of respective mappings. For each pair of AMs $m_1, m_2$ that cannot share the same ID set there shall be a row in matrix $A$ representing the inequality $x_{name(m_1)} + x_{name(m_2)} \leqslant 1$. $\mathbf{b}$ will be a vector of ones with corresponding length.

@@

\begin{define}[Heuristic Algorithm]
	A \textit{heuristic algorithm} is one that, in a reasonably short time, generates a good, maybe even optimal solution to an optimization problem. However, it won’t provide any formal guarantee about its quality.
\end{define}

@@

While a heuristic algorithm can be seen as a tool designed to solve one specific problem, the notion of a \textit{metaheuristic} reminds of a recipy to solve a whole family of problems. We shall be using the metaheuristic presented in Figure \ref{image-metaheuristic} to find optimal ID sets in this work. But before we start describing its structure and components, we need to introduce some more notions.

@@

\begin{define}[Solution Neighborhood]
	Neighborhood of a solution $\mathbf{x}$ in the solution space are all the other solutions close to $\mathbf{x}$ according to some metric.
\end{define}

The precise definition of the neighborhood is always adjusted according to specific needs. However, the neighborhood should be defined so that is \textit{continuous} (with respect to quality) to be useful. The exact reasons for this requirement are sketched in Section \ref{section-mip-ihs}.

@@

The notion of metaheuristic covers a wide range of topics in the field of heuristics, such as Tabu Search (see \cite{Glover:1997:TS:549765}), Ant Colony Optimization (see \cite{dorigo2004ant}) or Genetic Algorithms (see e.g. \cite{goldberg1989genetic}), to name a few.\\

@@

When we start to solve a problem using a metaheuristic approach, at first we have no solutions at all. The purpose of a construction heuristic (CH) is then to provide us with at least some solution. This may or may not be already the optimum, in the latter case it will be improved on later using improvement heuristics (IH). Some IHs can profit from a pool of several sub-optimal solutions, and some CHs can produce this pool from them.

@@

One of the most natural heuristics when dealing with the ID set problem can be described as follows: select from candidate AMs at random, if possible (addition would not violate the ID set condition) add them to the solution. This is obviously a hungry heuristic.

@@

This is again a hungry heuristic, the Listing \ref{listing-ch-fuzzy} contains its pseudocode.

@@

This trivial heuristics sorts all candidate AMs by their decreasing weights (see Section \ref{section-definitions-weight}) and then tries to iteratively add them to solution, if possible. This way it can create only one solution, and again, this is a hungry heuristic.

@@

This is basically a reversal of the idea from the \heu{Incremental} heuristic - start with a solution containing all the candidate AMs. This probably does not satisfy the ID set condition. Therefore, sort them by increasing size and start removing them from the solution, until it satisfies the ID set condition. Again, this is a hungry heuristic returning only one solution.

@@

To be able to create a pool of solutions, the GLPK input in MathProg langauge is always randomly shuffled by changing the order in which variables and constraints appear. This is causes the solver to explore the search tree in various orders, producing different solution in each of the time-constrained runs.

@@

Improvement heuristics in general start with a solution pool, attempt to improve one or more solutions in it and then return this improved pool in the end. We will need two notions to describe their behavior.

@@

is the attempt to move the solution towards the nearby local optimum in the solution space.

@@

is the attempt to move the solution away (escape) from from the local optimum, to be able to explore more of the solution space when the metaheuristic starts stagnating.\\

A metaheuristic needs to combine intensification and diversification tendencies to explore the solution space and at the same time arrive at a local optimum. Recall the requirement for the solution space to be continuous in terms of quality: this guarantees that as we approach a solution $\mathbf{x}$, the quality of solutions we encounter approaches the quality of $\mathbf{x}$.

@@

This ultimately trivial improvement heuristics does nothing. It simply returns the feasible pool unchanged. For the sake of completeness, see its Listing \ref{listing-ih-identity}.

@@

This trivial IH tries to improve the solution pool by removing the worst solution (i.e. the one with the lowest quality). This is interesting in cooperation with other improvement heuristics that increase the solution pool size, to keep it from growing by pruning inferior solutions.

@@

The number of AMs to remove from each solution is specified as fraction from $(0, 1)$ of the solution size (number of AMs in solution). For example, \heu{Random Remove} with $fraction = 0.1$ would remove 1 random AM from a solution containing 10 AMs and 2 from a solution containing 17 AMs (due to rounding).

This heuristic returns a pool of solutions of the same size as it got on its input.

@@

\heu{Mutation} is based on the following idea. We assume that an incumbent solution may already contain some AMs belonging to the optimal solution. We will take a random guess and fix some of these AMs, i.e. we add new constraints to the MIP formulation fixing values of the respective variables to 1.

@@

This improvement heuristic expands on the idea of \heu{Mutation}. But, instead of randomly selecting AMs in the incumbent solution, it looks for commonalities among the solutions in the pool. This is based on the hope that if more solutions agree on the same AMs, those are probably included in the optimal solution too.

\heu{Crossover} takes a parameter - fraction of solutions in the pool among which to look for commonalities. AMs found in every one of them are fixed in the modified MIP formulation the same way as in \heu{Mutation}. This again amounts to an intensification tendency.

Listing \ref{listing-ih-crossover} captures the process. \heu{Crossover} requires at least one solution in the pool, but to work properly, more are needed. Solutions are picked at random from the pool, common AMs found and fixed. GLPK is run again (with a time constraint) and the improved solution is added to the result pool.

@@

See Listing \ref{listing-ih-localbranching} for pseudocode. The heuristic requires a pool containing at least one solution, solves the modified MIP formulation using GLPK limited to some time again and adds the improved solution to the result pool.\\

@@

Now the IDREF set contains all the AMs whose images are a subset of the ID set image.

>>>>

In this chapter we introduce a new approach to finding maximum ID sets. First, we transform the problem formulation to maximum weighted IS problem formulation. Then we transform this into an MIP formulation, and demonstrate how this can be solved using a solver such as GLPK \cite{glpk}. \nomenclature{GLPK}{GNU Linear Programming Kit}
We will continue by~applying heuristic approaches to improve the performance of the process.

@@

An advantage of using \textit{Branch \& Bound} is~that while traversing the search tree it finds intermittent, sub-optimal solutions. It is thus possible to limit the~total search time and instead of the optimum take the best solution found so far.\\

We will now demonstrate the full process of~finding the optimal ID set of~an~example XML file using GLPK.

@@

The first three edges are due to the type collision ($y$), the last one is due to~$\iota(M_{y}^{a}) \cap \iota(M_{z}^{a}) = \{1\}$. The graph $G$ constructed in this way is shown in~Figure~\ref{image-mip-is-graph}.

\begin{figure}
  \centering
	\includegraphics[width=.25\textwidth]{images/is-representation}
  \caption{IS Representation Graph}
  \label{image-mip-is-graph}
\end{figure}

The next step is the MIP formulation. We do not need to translate from the IS formulation, as the translation from ID set formulation is straightforward, too. For each AM $m$ there will be one binary variable $x_{name(m)}$. Objective function coefficients in vector $\mathbf{c}$ will be weights of respective mappings. For each pair of~AMs $m_1, m_2$ that cannot share the same ID set there shall be a row in matrix $A$ representing the inequality $x_{name(m_1)} + x_{name(m_2)} \leqslant 1$. $\mathbf{b}$ will be a vector of~ones of corresponding length.

@@

\begin{define}[Heuristic Algorithm]
	A \textit{heuristic algorithm} is one that, in a reasonably short time, generates a good, maybe even optimal solution to an optimization problem. However, it will not provide any formal guarantee about its quality.
\end{define}

@@

While a heuristic algorithm can be seen as a tool designed to solve one specific problem, the notion of a \textit{metaheuristic} reminds of a recipe to solve a whole family of problems. We shall be using the metaheuristic presented in Figure \ref{image-metaheuristic} to find optimal ID sets in this work. But before we start describing its structure and components, we need to introduce some more notions.

@@

\begin{define}[Solution Neighborhood]
	Neighborhood of a solution $\mathbf{x}$ in~the~solution space are all the other solutions close to~$\mathbf{x}$ according to~some metric.
\end{define}

The precise definition of the neighborhood is always adjusted according to~specific needs. However, the neighborhood should be defined so that is \textit{continuous} (with respect to quality) to be useful. The exact reasons for this requirement are sketched in Section \ref{section-mip-ihs}.

@@

The notion of metaheuristic covers a wide range of topics in the field of~heuristics, such as Tabu Search (see \cite{Glover:1997:TS:549765}), Ant Colony Optimization (see \cite{dorigo2004ant}) or Genetic Algorithms (see e.g. \cite{goldberg1989genetic}), to name a few.\\

@@

When we start to solve a problem using a metaheuristic approach, at first we have no solutions at all. The purpose of a construction heuristic (CH) is then to provide us with at least some solution. This may or may not be already the~optimum, in the latter case it will be improved on later using improvement heuristics (IH). Some IHs can profit from a pool of several sub-optimal solutions, and some CHs can produce this pool from them.

@@

One of the most natural heuristics when dealing with the ID set problem can be described as follows: select from candidate AMs at random, if possible (addition would not violate the ID set condition) add them to the solution. This is obviously a greedy heuristic.

@@

This is again a greedy heuristic, the Listing \ref{listing-ch-fuzzy} contains its pseudocode.

@@

This trivial heuristic sorts all candidate AMs by their decreasing weights (see Section \ref{section-definitions-weight}) and then tries to iteratively add them to solution, if possible. This way it can create only one solution, and again, this is a greedy heuristic.

@@

This is basically a reversal of the idea from the \heu{Incremental} heuristic - start with a solution containing all the candidate AMs. This probably does not satisfy the ID set condition. Therefore, sort them by increasing size and start removing them from the solution, until it satisfies the ID set condition. Again, this is a greedy heuristic returning only one solution.

@@

To be able to create a pool of solutions, the GLPK input in MathProg langauge is always randomly shuffled by changing the order in which variables and constraints appear. This is causes the solver to explore the search tree in~various orders, producing different solution in each of the time-constrained runs.

@@

Improvement heuristics in general start with a solution pool, attempt to improve one or more solutions in it and then return this improved pool in~the~end. We will need two notions to describe their behavior.

@@

is the attempt to move the solution towards the nearby local optimum in~the~solution space.

@@

is the attempt to move the solution away (escape) from the~local optimum, to be able to explore more of the solution space when the metaheuristic starts stagnating.\\

A metaheuristic needs to combine intensification and diversification tendencies to explore the solution space and at the same time arrive at a local optimum. Recall the requirement for the solution space to be continuous in~terms of quality: this guarantees that as we approach a solution $\mathbf{x}$, the quality of solutions we encounter approaches the quality of $\mathbf{x}$.

@@

This ultimately trivial improvement heuristic does nothing. It simply returns the feasible pool unchanged. For the sake of completeness, see its Listing \ref{listing-ih-identity}.

@@

This trivial IH tries to improve the solution pool by removing the worst solution (i.e. the one with the lowest quality). This is interesting in cooperation with other improvement heuristics that increase the solution pool size, to keep it~from growing by pruning inferior solutions.

@@

The number of AMs to remove from each solution is specified as fraction from $(0, 1)$ of the solution size (number of AMs in solution). For example, \heu{Random Remove} with $fraction = 0.1$ would remove 1 random AM from a solution containing 10 AMs and 2 from a solution containing 17 AMs (due to~rounding).

This heuristic returns a pool of solutions of the same size as it got on input.

@@

\heu{Mutation} is based on the following idea. We assume that an incumbent solution may already contain some AMs belonging to the optimal solution. We will take a random guess and fix some of these AMs, i.e. we add new constraints to~the~MIP formulation fixing values of the respective variables to 1.

@@

This improvement heuristic expands on the idea of \heu{Mutation}. But, instead of~randomly selecting AMs in the incumbent solution, it looks for commonalities among the solutions in the pool. This is based on the hope that if more solutions agree on the same AMs, those are probably included in the optimal solution too.

\heu{Crossover} takes a parameter - fraction of solutions in the pool among which to look for commonalities. AMs found in every one of them are fixed in the~modified MIP formulation the same way as in \heu{Mutation}. This again amounts to~an~intensification tendency.

Listing \ref{listing-ih-crossover} captures the process. \heu{Crossover} requires at least one solution in~the pool, but to work properly, more are needed. Solutions are picked at~random from the pool, common AMs found and fixed. GLPK is run again (with a~time constraint) and the improved solution is~added to~the~result pool.

@@

See Listing \ref{listing-ih-localbranching} for pseudocode. The heuristic requires a pool containing at~least one solution, solves the modified MIP formulation using GLPK limited to~some time again and adds the improved solution to the result pool.\\

@@

Now the IDREF set contains all the AMs whose images are a subset of the~ID set image.
